Blockchain System

### 2021-01-27 papers/execution-model/model.tex
another invariant
some typing rules
some changes to typing

Example pseudo code
\begin{align*}
  \OPH &=  \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR \\
       & \MATCH\ {status\ \OPH}\ \WITH \\
       & \mid \STATUSPENDING \to \textrm{not yet included, continue waiting} \\
       & \mid \STATUSINCLUDING (n) \to \textrm{included with $n$
         blocks confirmation}\\
       & \mid \STATUSTIMEOUT \to \textrm{lost, try again}
\end{align*}
Grammar of types
\begin{align*}
  \TYPE & ::= \TADDR \mid \TOPH \mid \TPUH \mid \TPUK \mid \TCODE\ \TYPE \mid
          \TINT \mid \TUNIT \mid \TBOOL \mid \TSTRING \mid \TSTATUS \\
  & \mid \TYPE\to\TYPE \mid \TPAIR\ \TYPE\ \TYPE \mid \TLIST\ \TYPE
    \mid \TSUM\ \TYPE\ \TYPE \mid \TOPTION\ \TYPE \mid \TEXCEPTION
\end{align*}

@@

 \item $ \DOM (\CONTRACTORS) \subseteq \DOM (\CONTRACTORS')$
   \begin{itemize}
   \item for all $\PUH \in \DOM (\CONTRACTORS)$,
     $\CONTRACTORS (\PUH).\CODE = \CONTRACTORS' (\PUH).\CODE$
   \end{itemize}
\end{itemize}

@@

Define $ \Delta \vdash \CONTRACTORS$ by
\begin{itemize}
\item $\DOM (\Delta) = \DOM (\CONTRACTORS)$
\item $\CONTRACTORS (\PUH).\CODE : \Delta (\PUH) \to FIXEDOUTPUTTYPE$
\item $\CONTRACTORS (\PUH).\STORAGE : \TYPE_s$ where $\Delta (\PUH) =
  \TPAIR\ \TYPE_p\ \TYPE_s$. 
\end{itemize}

>>>>

Example pseudo code
\begin{align*}
  \OPH &=  \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR \\
       & \MATCH\ {status\ \OPH}\ \WITH \\
       & \mid \STATUSPENDING \to \textrm{not yet included, continue waiting} \\
       & \mid \STATUSINCLUDING (n) \to \textrm{included with $n$
         blocks confirmation}\\
       & \mid \STATUSTIMEOUT \to \textrm{lost, try again}
\end{align*}
Grammar of types
\begin{align*}
  \TYPE & ::= \TADDR \mid \TOPH \mid \TPUH\ \TYPE\ \TYPE \mid \TPUK \mid \TCODE\
          \TYPE\ \TYPE \mid \TTEZ \mid
          \TINT \mid \TUNIT \mid \TBOOL \mid \TSTRING \mid \TSTATUS \\
  & \mid \TYPE\to\TYPE \mid \TPAIR\ \TYPE\ \TYPE \mid \TLIST\ \TYPE
    \mid \TSUM\ \TYPE\ \TYPE \mid \TOPTION\ \TYPE \mid \TEXCEPTION
\end{align*}

Typing rules, relying on an external typing judgment
$\JTypeCode\CODE\TYPE$ for the contract language and
$\JTypeValue\STRING\TYPE$ for serialized values as stored on the blockchain.
\begin{mathpar}
  \inferrule{}{\JTypeExpr\TEnv\INT\TINT}

  \inferrule{}{\JTypeExpr\TEnv\OPH\TOPH}

  \inferrule{}{\JTypeExpr\TEnv\PUH{\TPUH\ ?\ ?}}

  \inferrule{}{\JTypeExpr\TEnv\PUK\TPUK}

  \inferrule{\JTypeCode \CODE{ \TPAIR\ \TYPE_p\ \TYPE_s}}{\JTypeExpr\TEnv\CODE{\TCODE\ \TYPE_p\ \TYPE_s}}

  \inferrule{}{\JTypeExpr\TEnv\NTEZ\TTEZ}

  \inferrule{}{\JTypeExpr\TEnv\SUNIT\TUNIT}

  \inferrule{}{\JTypeExpr\TEnv\FALSE\TBOOL}

  \inferrule{}{\JTypeExpr\TEnv\TRUE\TBOOL}

  \inferrule{}{\JTypeExpr\TEnv\STATUSPENDING\TSTATUS}

  \inferrule{}{\JTypeExpr\TEnv\STATUSTIMEOUT\TSTATUS}

  \inferrule{}{\JTypeExpr\TEnv{\STATUSINCLUDING (\INT)}\TSTATUS}

  \inferrule{}{\JTypeExpr\TEnv\ERROR\TEXCEPTION}

  \inferrule{}{\JTypeExpr\TEnv\VARIABLE{\TEnv (\VARIABLE)}}

  \inferrule{
    \JTypeExpr{\TEnv, \VARIABLE:\TYPE'}\EXPR{\TYPE}
  }{
    \JTypeExpr\TEnv{\lambda\VARIABLE.\EXPR}{\TYPE'\to\TYPE}}

  \inferrule{
    \JTypeExpr\TEnv\EXPR{\TYPE'\to\TYPE} \\
    \JTypeExpr\TEnv{\EXPR'}{\TYPE'}
  }{
    \JTypeExpr\TEnv{\EXPR\ \EXPR'}\TYPE
  }

  \inferrule{
    \JTypeExpr\TEnv\EXPR\TYPE \\
    \JTypeExpr\TEnv{\EXPR'}{\TYPE'}
  }{
    \JTypeExpr\TEnv{(\EXPR, \EXPR')}{\TPAIR\ \TYPE\ \TYPE'}
  }

  \inferrule{
    \JTypeExpr\TEnv{\EXPR}{\TPAIR\ \TYPE\ \TYPE'}
  }{
    \JTypeExpr\TEnv{\FST\ \EXPR}\TYPE
  }

  \inferrule{
    \JTypeExpr\TEnv{\EXPR}{\TPAIR\ \TYPE\ \TYPE'}
  }{
    \JTypeExpr\TEnv{\SND\ \EXPR}{\TYPE'}
  }
  \\\\
  \inferrule{
    \JTypeExpr\TEnv\EXPR\TEXCEPTION
  }{
    \JTypeExpr\TEnv{\RAISE\ \EXPR}\TYPE
  }

  \inferrule{
    \JTypeExpr\TEnv\EXPR\TYPE \\
    \JTypeExpr\TEnv{\EXPR'}{\TEXCEPTION\to\TYPE}
  }{
    \JTypeExpr\TEnv{\TRY\ \EXPR\ \EXCEPT\ \EXPR'}\TYPE
  }
  \\\\
  % too generic
  % \inferrule{
  %   \JTypeExpr\TEnv{\EXPR_1}\TTEZ \\
  %   \JTypeExpr\TEnv{\EXPR_2}\TPUK \\
  %   \JTypeExpr\TEnv{\EXPR_3}\TADDR \\
  %   \JTypeExpr\TEnv{\EXPR_4}\TSTRING \\
  %   \JTypeExpr\TEnv{\EXPR_5}\TTEZ
  % }{
  %   \JTypeExpr\TEnv{\TRANSFER[\EXPR_4]{\EXPR_1}{\EXPR_2}{\EXPR_3}{\EXPR_5}}\TOPH
  % }
  %   
  \inferrule{
    \JTypeExpr\TEnv{\EXPR_1}\TTEZ \\
    \JTypeExpr\TEnv{\EXPR_2}\TPUK \\
    \JTypeExpr\TEnv{\EXPR_3}\TPUK \\
    % \JTypeExpr\TEnv{\EXPR_4}\TSTRING \\
    \JTypeExpr\TEnv{\EXPR_5}\TTEZ
  }{
    \JTypeExpr\TEnv{\TRANSFER{\EXPR_1}{\EXPR_2}{\EXPR_3}{\EXPR_5}}\TOPH
  }

  \inferrule{
    \JTypeExpr\TEnv{\EXPR_1}\TTEZ \\
    \JTypeExpr\TEnv{\EXPR_2}\TPUK \\
    \JTypeExpr\TEnv{\EXPR_3}\TPUH\ \TYPE_p\ \TYPE_s \\
    \JTypeExpr\TEnv{\EXPR_4}{\TYPE_p} \\
    \JTypeExpr\TEnv{\EXPR_5}\TTEZ
  }{
    \JTypeExpr\TEnv{\TRANSFER[\EXPR_4]{\EXPR_1}{\EXPR_2}{\EXPR_3}{\EXPR_5}}\TOPH
  }

  \inferrule{
    \JTypeExpr\TEnv{\EXPR_1}\TTEZ \\
    \JTypeExpr\TEnv{\EXPR_2}\TPUK \\
    \JTypeExpr\TEnv{\EXPR_3}{\TCODE\ (\TPAIR\ \TYPE_p\ \TYPE_s)} \\
    \JTypeExpr\TEnv{\EXPR_4}\TYPE_s \\ % really \TSTRING \\
    \JTypeExpr\TEnv{\EXPR_5}\TTEZ
  }{
    \JTypeExpr\TEnv{\ORIGINATE{\EXPR_1}{\EXPR_2}{\EXPR_3}{\EXPR_4}{\EXPR_5}}{
      \TPUH\ \TYPE_p\ \TYPE_s} 
  }
\end{mathpar}

@@

 \item $ \DOM (\CONTRACTORS) \subseteq \DOM (\CONTRACTORS')$
   \begin{itemize}
   \item for all $\PUH \in \DOM (\CONTRACTORS)$,
     $\CONTRACTORS (\PUH).\CODE = \CONTRACTORS' (\PUH).\CODE$
   \end{itemize}
 \item invariant for contractors:
   for all $\PUH \in \DOM (\CONTRACTORS)$,
   $\CONTRACTORS (\PUH) = \Angle{ \CODE, \TIME, \BAL, \STORAGE}$
   implies that $\PUH = \GENERATEHASH (\CODE, \TIME)$.
\end{itemize}

@@

Define $ \Delta \vdash \CONTRACTORS$ by
\begin{itemize}
\item $\DOM (\Delta) = \DOM (\CONTRACTORS)$
\item $\JTypeCode{\CONTRACTORS (\PUH).\CODE} {\Delta (\PUH)}$
\item $\JTypeValue{\CONTRACTORS (\PUH).\STORAGE}{ \TYPE_s}$ where
  $\Delta (\PUH) = \TPAIR\ \TYPE_p\ \TYPE_s$. 
\end{itemize}
